---
output: html_document
editor_options: 
  chunk_output_type: console
---
Red Wine Quality Analysis
========================================================


The dataset was created using red wine samples.
The inputs include objective tests (e.g. PH values) and the output is based on 
sensory data (median of at least 3 evaluations made by wine experts). 
Each expert graded the wine quality between 0 (very bad) and 
10 (very excellent). 

The dataset is related to red variants of the Portuguese "Vinho Verde" wine.
Due to privacy and logistic issues, only physicochemical (inputs) and 
sensory (the output) variables are available (e.g. there is no data about 
grape types, wine brand, wine selling price, etc.).

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(dplyr)
library(RColorBrewer)
library(scales)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityReds.csv')
str(wine)
wine$X <- factor(wine$X)
rownames(wine) <- wine$X
summary(wine)
```



# Univariate Plots Section


In this section we will look at different variables individually

## Quality

```{r echo=FALSE, warning=FALSE, Univariate_Plots}
summary(wine$quality)
ggplot(aes(x=quality), data=wine)+
  geom_histogram(binwidth = 1, fill='blue', color='black')+
  ggtitle("Quality counts")
```

Here we see in this graph that we have more observations for the normal wine
than for poor and good quality.

## Fixed Acidity (Nonvolatile acids)

```{r echo=FALSE, warning=FALSE}
summary(wine$fixed.acidity)
ggplot(aes(x=fixed.acidity), data=wine)+
  geom_histogram(binwidth = 0.25, fill='green', color='black')+
  ggtitle("Histogram for fixed acidity count")

```

The data is skewed. The values are more concentrated towards the left. With a 
few outliers near 15 and 16.

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=fixed.acidity), data=wine)+
  geom_histogram(binwidth = 0.25, fill='green', color='black')+
  xlim(min(wine$fixed.acidity), quantile(wine$fixed.acidity, 0.99))+
  ggtitle("Histogram for fixed acidity count for (0-99) percentile")
```

In this graph we removed the upper 1% of values. and now the graphed is less 
skewed.

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=fixed.acidity), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  scale_x_log10()+
  ggtitle("Histogram for fixed acidity (log10) count")
```

This graphs is for log10 tranformation of data and seem more normal without 
removing the outliers.
We have max values between 7 and 8.5 approx and then on max value at 10.

## Volatile acidity (Acetic acid)

```{r echo=FALSE, warning=FALSE }
summary(wine$volatile.acidity)
ggplot(aes(x=volatile.acidity), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  ggtitle("Histogram for volatile acidity count")
```

This is the histogram for volatile acidity count. We can see a few outliers on 
both tails, with min val of 0.12 and max of 1.58 
There are gaps on both end and these values seems to be outliers, specially on 
the right tail.

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=volatile.acidity), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  xlim(min(wine$volatile.acidity), quantile(wine$volatile.acidity, 0.99))+
  ggtitle("Histogram for volatile acidity count")
```

After removing the upper 1% values graph is more clear with some spikes at
around 0.40, 0.50 and 0.60

## Citric acid

```{r echo=FALSE, warning=FALSE }
summary(wine$citric.acid)
ggplot(aes(x=citric.acid), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  ggtitle("Histogram for citric acid count")
```

This is interesting we have values between 0 and 1. with max values at 0 and 
then we see a high number for 0.50 and a very slight rise at 0.25
We see that 75% is at 0.42, that is less than half of the range of values.

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=citric.acid), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  xlim(0, quantile(wine$citric.acid, 0.99))+
  ggtitle("Histogram for citric acid count (0-99) percentile")
```

## Residual sugar

```{r echo=FALSE, warning=FALSE }
summary(wine$residual.sugar)
ggplot(aes(x=residual.sugar), data=wine)+
  geom_histogram(binwidth = 0.05, fill='green', color='black')+
  ggtitle("Histogram for residual sugar count")
```

The data is very much normal with a long tail towards the right

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=residual.sugar), data=wine)+
  geom_histogram(binwidth = 0.05, fill='green', color='black')+
  scale_x_sqrt()+
  ggtitle("Histogram for residual sugar count (sqrt)")
```

With sqrt it is still concentrated to the left

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=residual.sugar), data=wine)+
  geom_histogram(binwidth = 0.05, fill='green', color='black')+
  scale_x_log10()+
  ggtitle("Histogram for residual sugar count (log10)")
```

We can see with log10 transformation that we have high count for approx 2.2

## Chlorides (Salt)

```{r echo=FALSE, warning=FALSE }
summary(wine$chlorides)
ggplot(aes(x=chlorides), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  ggtitle("Histogram for chlorides count")
```

this data has similar shape to sugar residue data

```{r echo=FALSE, warning=FALSE }
ggplot(aes(x=chlorides), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  scale_x_log10(breaks = seq(0.01, 0.6, 0.01))+
  ggtitle("Histogram for chlorides count (log10)")
```

So we have a clear spike at 0.08 and very clear outliers near 0.01 and 0.6

## Sulfur dioxides and Sulphates

```{r echo=FALSE, warning=FALSE }
summary(wine$free.sulfur.dioxide)
summary(wine$total.sulfur.dioxide)
summary(wine$sulphates)
p1<-ggplot(aes(x=free.sulfur.dioxide), data=wine)+
  geom_histogram(binwidth = 1, fill='green', color='black')+
  ggtitle("Histogram for free sulfur dioxide count")
p2<-ggplot(aes(x=total.sulfur.dioxide), data=wine)+
  geom_histogram(binwidth = 5, fill='green', color='black')+
  ggtitle("Histogram for total sulfur dioxide count")
p3<-ggplot(aes(x=sulphates), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  scale_x_continuous(breaks = seq(2,4.5, 0.125))+
  ggtitle("Histogram for free sulphates count")
grid.arrange(p1,p2,p3,ncol=1)
```

```{r echo=FALSE, warning=FALSE }

cuberoot_trans <- function() {trans_new('cuberoot',
                                       transform = function(x) x^(1/3),
                                       inverse = function(x) x^(3))}
```

```{r echo=FALSE, warning=FALSE }
p1<-ggplot(aes(x=free.sulfur.dioxide), data=wine)+
  geom_histogram(binwidth = 0.1, fill='green', color='black')+
  scale_x_continuous(trans = cuberoot_trans(), breaks = seq(0,80, 5))+
  ggtitle("Histogram for free sulfur dioxide count (cuberoot)")
p2<-ggplot(aes(x=total.sulfur.dioxide), data=wine)+
  geom_histogram(binwidth = 0.05, fill='green', color='black')+
  scale_x_log10(breaks=seq(0,300, 25))+
  ggtitle("Histogram for total sulfur dioxide count (log10)")
p3<-ggplot(aes(x=sulphates), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  scale_x_log10(breaks = seq(2,4.5, 0.125))+
  ggtitle("Histogram for free sulphates count")
grid.arrange(p1,p2,p3,ncol=1)
```

In these graphs we have the cuberoot of free sulfur dioxide, log10 oftotal sulfur 
dioxide and again cuberoot of sulphates

## Density

```{r echo=FALSE, warning=FALSE }
summary(wine$density)
ggplot(aes(x=density), data=wine)+
  geom_histogram(binwidth = 0.0001, fill='green', color='black')+
  ggtitle("Histogram for free density count")
```

The density values are very small but the data seems normally distributed.

## pH

```{r echo=FALSE, warning=FALSE }
summary(wine$pH)
ggplot(aes(x=pH), data=wine)+
  geom_histogram(binwidth = 0.01, fill='green', color='black')+
  scale_x_continuous(breaks = seq(2,4.5, 0.125))+
  ggtitle("Histogram for free pH count")
```

Same goes for our pH index. 


# Univariate Analysis

### What is the structure of your dataset?

We have 1599 observations and 12 features in our data

All the test values are continuous, with varing ranges and the output quality is
a scale:
   0  ------->  10
(Worst)       (Best)

The dataset is biased as there are more observations for normal quality as for 
the good and poor ones.
Interesting thing is the scale for quality is 0-10 but the min-max values for
data are 3-8, no wine is very bad and none is exceptionally good.

### What is/are the main feature(s) of interest in your dataset?

The main feature is the quality of wine. I'd like to determine what determines 
the quality of a wine. 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The acids in our dataset and suger and salt(chloride) are the more likely
candidates for the wine quality from what I understand about wines.
the pH may depend upon the concentration of acids and the density on sugar 
content.

### Did you create any new variables from existing variables in the dataset?

No I didn't create new variables.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Most of the distributions were skewed with outliers towards max values.
I log10, sqrt and cuberoot transformed or removed upper 1% values to plot in
these distribuions. We know that we have very few good and poor wines. These 
outlier values may be related to these observations, which I would like to 
explore more in the next sections.

# Bivariate Plots Section

In this section we are going to look at pairs of features, how they are related
to one another.

## Fixed acidity and Citic Acid vs Quality

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(y=fixed.acidity, x=quality), data=wine)+
  geom_jitter(color='red')+
  scale_x_continuous(breaks=seq(0,10, 1))+
  ggtitle("Quanlity vs fixed acidity scatterplot")

```

In this acidity vs quality section we see that the acidic quantity is more in 
normal wine and we know that we have more observations for normal wines, which 
may produce this bias

```{r echo=FALSE, warning=FALSE}
p1<-ggplot(aes(y=fixed.acidity, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quanlity vs fixed acidity Mean and Median")
p2<-ggplot(aes(y=citric.acid, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quanlity vs citric acid Mean and Median")
grid.arrange(p1,p2,ncol=1)
```

Here we see that as the fixed acidity increses as the quality increases but 
after quality level 7 there is a fall in acidity.

In second graph it shows an upward trend in citric acid with the quality of 
wine, the more citric flavour the better the wine. This trend is similar to 
fixed acidity trend.

## Volatile acidity vs Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=volatile.acidity, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quanlity vs volatile acidity Mean and Median")
```

We see a downwards trend in acidity as the quality increases and at level 8 the 
meadian value is same as level same other values show a slight increase

## Residual sugar vs Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=residual.sugar, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quanlity vs residual sugar Mean and Median")
```

## Chlorides vs Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=chlorides, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quanlity vs chlorides Mean and Median")
```

## Sulfur dioxides vs Quality

```{r echo=FALSE, warning=FALSE}
p1<-ggplot(aes(y=free.sulfur.dioxide, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quality vs free sulfur dioxide Mean and Median")
p2<-ggplot(aes(y=total.sulfur.dioxide, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quality vs total sulfur dioxide Mean and Median")
grid.arrange(p1, p2, ncol=1)
```

We see a similar trend in both values with respect to qaulity because total 
sulfur dioxide is sum of both free and bound sulfur dioxides. For normal wines 
the value is high and for both poor and good wines the value is low

## Sulphates vs Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=sulphates, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quality vs sulphates Mean and Median")

```

This shows a positive relation between sulphates and quality

## pH vs Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=pH, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quality vs pH Mean and Median")
```

For poor wines the pH level is high but for good wines the pH level is low. 
Low pH levels cooresponds to more acidity. This represents the overall acidity 
of our wine. 

## Alcohol vs Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=alcohol, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Quality vs Alcohol Mean and Median")
```

This is again a upwards trend. The quality is getting better with the alcohol 
content

## Density vs quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(y=density, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks=seq(0,10, 1))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  ggtitle("Density vs Quality")
```

This shows a low density for good wines

## Acids, Sugar and salt

```{r echo=FALSE, warning=FALSE}
p1<-ggplot(aes(y=citric.acid, x=residual.sugar), data=wine)+
  geom_point(alpha=1/20, color='red',position = position_jitter(h=0)) +
  xlim(range=c(0, quantile(wine$residual.sugar, 0.99)))+
  ggtitle("Citric acid vs Sugar (0-99%)")
p2<-ggplot(aes(y=volatile.acidity, x=residual.sugar), data=wine)+
  geom_point(alpha=1/20, color='red', position = position_jitter(h=0)) +
  xlim(range=c(0, quantile(wine$residual.sugar, 0.99)))+
  ggtitle("Volatile acidit vs Sugar (0-99%)")
p3<-ggplot(aes(y=citric.acid, x=chlorides), data=wine)+
  geom_point(alpha=1/20, color='red', position = position_jitter(h=0)) +
  scale_x_log10()+
  ggtitle("Citric acid vs Chlorides (log10)")
p4<-ggplot(aes(y=volatile.acidity, x=chlorides), data=wine)+
  geom_point(alpha=1/20, color='red', position = position_jitter(h=0)) +
  scale_x_log10()+
  ggtitle("Volatile acidity vs Chloride (log10)s")
grid.arrange(p1,p2,p3,p4, ncol=2)

```

There seems to be no correlation between acids, sugar and salt


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The citric acid, sulphates and alcohol has a positive correlation with the 
quality of wine.

Volatile acids, density, residual sugar and chlorides have a negative 
correlation with the quality of wine.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

as the pH lowers the quality of wine increases, which means the wine becomes 
more acidic but we saw a negative correlation with volatile acids and positive 
for non volatile acids. this shows that the overall acidity of wine has a 
positive correlation with quality.

Most of the relationships are not directly linear, there is a behaviour 
change in relationship around level 7-8 or 3-4

### What was the strongest relationship you found?

Citric acid, volatile acid, Chlorides and sulphates are in strong relation with
quality of wine

# Multivariate Plots Section


## Citric acid vs Fixed acidity and Volatile acidity on quality

```{r echo=FALSE, Multivariate_Plots}
p1<-ggplot(aes(y=fixed.acidity, x=citric.acid), data=wine)+
  geom_point(aes(color=factor(quality)), size=2)+
  scale_color_brewer(type = 'div')+
  ggtitle("Citric acid vs Fixed acidity")
p2<-ggplot(aes(y=volatile.acidity, x=citric.acid), data=wine)+
  geom_point(aes(color=factor(quality)), size=2)+
  scale_color_brewer(type = 'div')+
  ggtitle("Citric acid vs Volatile acidity")
grid.arrange(p1,p2,ncol=1)
```

In the first graph it shows that the good quality wines have more fixed acidity
and more citric acid
And in second graph it shows good wines with low volatile and high cirtic acid
contents

## pH vs Alcohol on Quality

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=pH, y=alcohol), data=wine)+
  geom_point(aes(color=factor(quality)), size=2)+
  scale_color_brewer(type = 'div')+
  ggtitle("pH vs Alcohol acidity")
```

Here we see that for the same value of pH the quality of wine increases as the 
alcohol content increases

## Acids vs Alcohol

```{r echo=FALSE, warning=FALSE}
p1<-ggplot(aes(x=citric.acid, y=alcohol), data=wine)+
  geom_point(aes(color=factor(quality)), size=2)+
  scale_color_brewer(type = 'div')+
  ggtitle("citric acid vs Alcohol acidity")
p2<-ggplot(aes(x=volatile.acidity, y=alcohol), data=wine)+
  geom_point(aes(color=factor(quality)), size=2)+
  scale_color_brewer(type = 'div')+
  ggtitle("volatile acidity vs Alcohol acidity")
grid.arrange(p1, p2, ncol=1)
```

In both graphs for the same value for citric or volatile acids the quality of 
wine increases as alcohol content increases

## Sugar vs alcohol

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x=residual.sugar, y=alcohol), data=wine)+
  geom_point(aes(color=factor(quality)), size=2)+
  scale_x_log10()+
  scale_color_brewer(type = 'div')+
  ggtitle("Sugar vs Alcohol acidity")
```

Alcohol has same relation with sugar as with acids for the quality

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Great quality wines had high citric acid and fixed acids content and low 
volatile acids. And for the same pH level the alcohol has positive relation to 
wine quality.

### Were there any interesting or surprising interactions between features?

The alcohol content seems to have a very strong relationship with the wine 
quality for other variables constant the incerease in alcohol content increased
the quality of wine.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(y=sulphates, x=quality), data=wine)+
  geom_line(stat='summary', fun.y=mean, size=1, aes(color='Mean'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.5),
            size=1, aes(color='Median'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.25),
            size=1, aes(color='25%'))+
  geom_line(stat='summary', fun.y=quantile, fun.args=list(probs=0.75),
            size=1, aes(color='75%'))+
  scale_x_continuous(breaks = seq(0,10, 1))+
  scale_y_continuous(breaks = seq(0, 1, 0.025))+
  scale_colour_manual(values = c("Mean"="red", "Median"="green", 
                                 "25%"="blue", "75%"="blue")) +
  xlab("Quality (score between 0 and 10)")+
  ylab("Sulphates (potassium sulphate - g / dm3)")+
  ggtitle("Sulphates vs Quality (Mean, Median, 25 Percentile and 75 Percentile)")
```

### Description One

In this graph we see the sulphates content vs the quality of wine. We see that 
as the sulphate content in wine increases the quality of wine also increases.
the mean value shows a very linear relationship between the two values.


```{r echo=FALSE}
wine$Wine_Quality<-factor(wine$quality)
```

### Plot Two

```{r echo=FALSE, Plot_Two}
p1<-ggplot(aes(y=fixed.acidity, x=citric.acid), data=wine)+
  geom_point(aes(color=Wine_Quality), size=2)+
  scale_color_brewer(type = 'div')+
  xlab("Citric acid (g / dm^3)")+
  ylab("Fixed acid (tartaric acid - g / dm^3)")+
  ggtitle("Citric acid vs Fixed acid")
p2<-ggplot(aes(y=volatile.acidity, x=citric.acid), data=wine)+
  geom_point(aes(color=Wine_Quality), size=2)+
  scale_color_brewer(type = 'div')+
  xlab("Citric Acid (g / dm^3)")+
  ylab("Volatile acid (acetic acid - g / dm^3)")+
  ggtitle("Citric acid vs Volatile acid")
grid.arrange(p1,p2,ncol=1)
```

### Description Two

In the first plot we see that good wines have the high citric acid value and 
also high fixed acid value.

In the second graph we observe that for a better quality wine the citric acid 
has high content value but the lower volatile content. These two acids have 
opposite trends for the quality.

### Plot Three
```{r echo=FALSE, Plot_Three}
p1<-ggplot(aes(x=citric.acid, y=alcohol), data=wine)+
  geom_point(aes(color=Wine_Quality), size=2)+
  scale_color_brewer(type = 'div')+
  xlab("Citric acid (g / dm^3)")+
  ylab("Alcohal (% by volume)")+
  ggtitle("Citric acid vs Alcohol")
p2<-ggplot(aes(x=volatile.acidity, y=alcohol), data=wine)+
  geom_point(aes(color=Wine_Quality), size=2)+
  scale_color_brewer(type = 'div')+
  xlab("Volatile acid (acetic acid - g / dm^3)")+
  ylab("Alcohal (% by volume)")+
  ggtitle("Volatile acid vs Alcohol")
grid.arrange(p1, p2, ncol=1)
```

### Description Three

In both the plots we see the trends of previous plots. Also when the acid level 
is a constant the quality of wine increases as the alcohol content is 
increasing. This shows a strong positive correlation between alcohol and wine 
qulality.

------

# Reflection

The citric acid can add "freshness" and flavour to a wine and we saw that the 
quality of wine increases with the increase in citric acid content in wine.

Similarly, the high value of volatile acid (acetic acid) can lead to unpleasant,
vinegar taste and in our plot we observed that in high quality wine this acid 
content is low.

Also, the pH value is low in good wines, in other words, the acidity of wine is 
high in good quality wines.

In most cases, other values remains constant the increase in alcohol content 
increases the wine quality.

And, sugar, salt, density all have a negative relationships to wine quality

In the plots therse are too many outliers, as we know that the data is biased with 
more observations for normal wines, we may analyze those outlier values for the
wine quality.

------

# Citations

[1] P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
    Modeling wine preferences by data mining from physicochemical properties.
    In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.
    Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
                [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib